NN Portfolio Optimierung

### 2024-02-13 paper/Paper.tex
add pictures and equations to paper

            The updated dataset has 10,000 images (first-batch) that are all 32x32 and in JPG format. 
            Also, the 10 classes have been combined into one natural image class. 
            I have only selected 10,000 because there are only 7,753 suitable graphs in the scraped graph dataset. 
            Had I selected more that would have led to an imbalanced dataset in the model. 
            In my project, the CIFAR-10 dataset will be labeled as CIFAR or natural.

@@

            In the end, four classes were kept, bar chart, diagram, graph, and pie chart, 
            so the total images now used were 7,753. 
            he images came in various dimensions and were resized to the natural images 32x32 and converted to JPG format. 
            I’ll note here that JPG was chosen because it only has 3 color channels and no alpha channel. 
            If PNG is used then it could have a shape of (32,32,4) which adds unnecessary parameters for the scope of my objectives.

@@

            Early on, I pre-identified possible misclassifications once I scanned through the set of generated graphs. 
            The similarities between some of the graphs, especially the log-normal and normal distribution, were considerable. 
            They were not ‘duds’ like previously mentioned -- they looked typical, but because of the density functions’ parameters, 
            they were similar in appearance to a point where it would become a 50-50 decision if I had to classify them myself. 
            The cases when this happened are a minority, but it was important to keep this in mind as I evaluated the models later.

            Normal Distribution Density Function

            Log-Normal Distribution Density Function

            Exponential Distribution Density Function

            Uniform Distribution Density Function

@@

            \begin{figure}[htp]

                    \includegraphics[scale=0.5]{types-of-graphs.png}
                    \caption{Feed-Forward Neuronales Netz \cite{bishop1995}}

@@

            The ANN process used in my model is called a feed-forward neural network or FNN. 
            The structure of the model consists of three parts: 
            an input layer, one or many hidden layers consisting of neurons, and an output layer. 
            The two main processes in the model are forward-propagation and back-propagation, 
            and can be summed up by their respective goals, producing an estimate and minimizing the loss. 
            Forward-propagation begins once the model is built and the data is split into the input values x and the target values y. 
            In this scenario the goal is to use x to predict y.

@@

                    \includegraphics[scale=0.5]{types-of-graphs.png}
                    \caption{Long Short-Term Memory Unit \cite{yan2016}}

@@

@@

                    \includegraphics[scale=0.5]{types-of-graphs.png}
                    \caption{Long Short-Term Memory Unit \cite{data-basecamp}}

@@

                The most important part of a CNN is its convolutional layers 
                since they are the means in which the original image gets ‘boiled down’ into its ‘essence’. 
                I use this analogy because one can imagine a set of ingredients, pixels making up an image, 
                that get thrown into a boiling pot where the water evaporates leaving only the important flavors behind. 
                This visceral example can be summed up more literally, 
                where the original image gets transformed layer by layer from the original pixel values to the final class scores. 
                By the end, the full image will be reduced into a single vector of class scores, arranged along the depth dimension. 
                What are these class scores? They indicate the estimate for each class or y target. 
                As an example, the final output layer for CIFAR-10 would have dimensions of (1, 1, 10) 
                since it has a total of 10 classes that an image can be sorted into \cite{cs231n2023} -- airplane, dog, cat, etc.

@@

                Output Volume Size Equation

            % \subsubsection{Max Pooling Layers}

@@

                Max Pooling Function

@@

                As stated earlier, the activation function simply tells the neuron to ‘fire’ or ‘rest’ 
                which in the case of the activation function I used, rectified linear unit function or ReLU, 
                the value returned is either 0 if x < 0 or x if x > 0 \cite{agarap2019}. 
                ReLU is used as an activation function for a few key reasons. First, the simple nature of the function, 
                f(x) = max(0,x), eliminates complex calculations and reduces the processing demands. 
                This is especially important in convolutional layers where the amount of parameters can skyrocket 
                and increase the time needed to learn. This works in practice by promoting sparsity in the model 
                and sparsity refers to the scenario where most of the cell entries in a matrix, 
                or in this case tensor, are zero \cite{giskard}.

                Figure of the Rectified Linear Unit \cite{agarap2019}

                \begin{multicols}{2}

                    \begin{Large} \[ Sharpe = \frac{R_p - R_f}{\sigma_p} \] \end{Large}

                    \vfill

                    \begin{small}

                        \noindent $R_p$ = Portfolio Renditen \newline 
                        $R_f$ = risikofreier Zinssatz \newline 
                        $\sigma_p$ = Standardabweichung der Renditen

                    \end{small}

                \end{multicols}

@@

                The next important topic to cover is the loss function used since the learning process can only happen 
                if the model has a loss function to calculate the error. 
                The loss function in the models I built was the cross-entropy loss function. 
                These functions, used in classification problems, calculate how accurate the machine learning 
                or deep learning model is by defining the difference between the estimated probability 
                with our desired outcome \cite{365team2023}. These functions come in two versions, 
                one for binary classification and one for categorical classification. 
                My model uses categorical cross-entropy because it is a multi-class classification problem, 
                which means there is only one correct class per x input or image.

                Binary Cross-Entropy Function

                Categorical Cross-Entropy Function

>>>>

            The updated dataset has 10,000 images (first-batch) that are all 32x32 and in JPG format. 
            Also, the 10 classes have been combined into one natural image class. 
            I have only selected 10,000 because there are only 7,753 suitable graphs in the scraped graph dataset. 
            Had I selected more that would have led to an imbalanced dataset in the model. 
            In my project, the CIFAR-10 dataset will be labeled as CIFAR or natural. See Figure \ref{cifar-overview} for examples.

                    \includegraphics[scale=0.6]{natural_overview.png}
                    \caption{CIFAR-10 Natural Images (CIFAR) \cite{krizhevsky2009}}
                    \label{cifar-overview}

@@

            In the end, four classes were kept, bar chart, diagram, graph, and pie chart, 
            so the total images now used were 7,753. 
            he images came in various dimensions and were resized to the natural images 32x32 and converted to JPG format. 
            I’ll note here that JPG was chosen because it only has 3 color channels and no alpha channel. 
            If PNG is used then it could have a shape of (32,32,4) which adds unnecessary parameters for the scope of my objectives.
            See Figure \ref{scraped-overview} for examples.

                    \includegraphics[scale=0.6]{scraped_overview_original.png}
                    \caption{Scraped Graphs (SCP) \cite{sunedition2021}}
                    \label{scraped-overview}

@@

            Early on, I pre-identified possible misclassifications once I scanned through the set of generated graphs. 
            The similarities between some of the graphs, especially the log-normal and normal distribution, were considerable. 
            They were not ‘duds’ like previously mentioned -- they looked typical, but because of the density functions’ parameters, 
            they were similar in appearance to a point where it would become a 50-50 decision if I had to classify them myself. 
            The cases when this happened are a minority, but it was important to keep this in mind as I evaluated the models later.
            See Figure \ref{generated-overview} for examples. The density functions that created the random variables in the graph
            are listed below.

                    \includegraphics[scale=0.6]{norm_460x345_overview.png}
                    \caption{Generated Graphs (GEN)}
                    \label{generated-overview}

            \textbf{Normal Distribution Density Function  $ X \sim \mathcal{N}(\mu^2, \sigma)  $}

            \begin{Large}

                \[ f_X(x) = \frac{1}{\sqrt{2\pi \sigma^2}} exp\left[{-\frac{(x - \mu)^2}{2 \sigma^2}}\right] \]

            \end{Large}

            \textbf{Log-Normal Distribution Density Function  $ X \sim L\mathcal{N}(\mu^2, \sigma) $}

            \begin{Large}

                \[ f_X(x) = \frac{1}{\sqrt{2\pi \sigma^2}} exp\left[{-\frac{(log(x) - \mu)^2}{2 \sigma^2}}\right] \]

            \end{Large}

            \textbf{Exponential Distribution Density Function  $ X \sim \textnormal{Exp}(\lambda) $}

            \begin{Large}

                \[ f_X(x) = \lambda exp(- \lambda x) \]

            \end{Large}

            \textbf{Uniform Distribution Density Function  $ X \sim \textnormal{U}[a, b] $}

            \begin{Large}

                \[ f_X(x) = \frac{1}{b - a} \]

            \end{Large}

@@

@@

            The ANN process used in my model is called a feed-forward neural network or FNN. 
            The structure of the model consists of three parts and are shown in Figure \ref{nn-bishop}: 
            an input layer, one or many hidden layers consisting of neurons, and an output layer. 
            The two main processes in the model are forward-propagation and back-propagation, 
            and can be summed up by their respective goals, producing an estimate and minimizing the loss. 
            Forward-propagation begins once the model is built and the data is split into the input values x and the target values y. 
            In this scenario the goal is to use x to predict y.

@@

                    \includegraphics[scale=1]{neural-network-bishop.png}
                    \caption{Neural Network \cite{bishop1995}}
                    \label{nn-bishop}

@@

                The most important parts of a CNN are its convolutional layers and pooling layers 
                since they are the means in which the original image gets ‘boiled down’ into its ‘essence’.
                I use this analogy because one can imagine a set of ingredients, pixels making up an image, 
                that get thrown into a boiling pot where the water evaporates leaving only the important flavors behind. 
                This visceral example can be summed up more literally, 
                where the original image gets transformed layer by layer from the original pixel values to the final class scores. 
                By the end, the full image will be reduced into a single vector of class scores, arranged along the depth dimension. 
                What are these class scores? They indicate the estimate for each class or y target. 
                As an example, the final output layer for CIFAR-10 would have dimensions of (1, 1, 10) 
                since it has a total of 10 classes that an image can be sorted into \cite{cs231n2023} -- airplane, dog, cat, etc.
                The step by step process can be seen more clearly in Figure \ref{cnn-overview-fig},
                where the input image of car will be transformed into the most important features and then classified as a car.

@@

                        \includegraphics[scale=0.65]{cnn-overview.png}
                        \caption{Long Short-Term Memory Unit \cite{cs231n2023}}
                        \label{cnn-overview-fig}

@@

@@

                \textbf{Output Volume Size Equation:}

                \begin{large}

                    \[ \textnormal{V} = \frac{W - F + 2P}{S} + 1    \]

                \end{large}

                Where W := input volume size, F := receptive field size (kernel), P := amount of zero padding, and S := stride of kernel

@@

                \textbf{Max Pooling Function:}

                \begin{large}

                    \[ \textnormal{M} = max(f_i)  \]

                \end{large}

                for i = 1, ... , F feature map patches

                        \includegraphics[scale=0.9]{maxpooling.png}
                        \caption{Maxpooling Example \cite{kumar2021} Credit: Codicals}
                        \label{maxpool-fig}

@@

                As stated earlier, the activation function simply tells the neuron to ‘fire’ or ‘rest’ 
                which in the case of the activation function I used, rectified linear unit function or ReLU, 
                the value returned is either 0 if $ x < 0 $ or x if $ x > 0 $ \cite{agarap2019}. 
                ReLU is used as an activation function for a few key reasons. First, the simple nature of the function, 
                f(x) = max(0,x), eliminates complex calculations and reduces the processing demands. 
                This is especially important in convolutional layers where the amount of parameters can skyrocket 
                and increase the time needed to learn. This works in practice by promoting sparsity in the model 
                and sparsity refers to the scenario where most of the cell entries in a matrix, 
                or in this case tensor, are zero \cite{giskard}.

                \textbf{Rectified Linear Units (ReLU):}

                \begin{large}

                    \[ f(x) = x^+ = max(0, x) = \frac{x + \vert x \vert}{2} = 
                    \left \{^{\textnormal{x if $x > 0$}}_{\textnormal{0 otherwise}} \right. \]

                \end{large}

                        \includegraphics[scale=0.8]{relu.png}
                        \caption{Rectified Linear Unit \cite{agarap2019}}
                        \label{relu-fig}

@@

                The next important topic to cover is the loss function used since the learning process can only happen 
                if the model has a loss function to calculate the error. 
                The loss function in the models I built was the cross-entropy loss function. 
                These functions, used in classification problems, calculate how accurate the machine learning 
                or deep learning model is by defining the difference between the estimated probability 
                with the desired outcome \cite{365team2023}. These functions come in two versions, 
                one for binary classification and one for categorical classification. 
                My model uses categorical cross-entropy because it is a multi-class classification problem, 
                which means there is only one correct class per x input or image. Had I evaluated if multiple graphs were in an image,
                then it would become a multi-label classification problem, so the final prediction vector would have
                a dimensionality of the number of classes, but each element in the vector could be 0 or 1 depending 
                if the image contains two classes. For example, if an image had two graphs with one uniform distribution 
                and one normal distribution, then it would look like [0, 0, 1, 1] (classes sorted alphabetically) \cite{gomez2018}.

                \textbf{Binary Cross-Entropy Function:}

                \begin{large}

                    \[ BCEL = - \frac{1}{N} \sum_{i=1}^N (y_i \cdot log(p_i) + (1 - y_i) \cdot log(1-p_i)) \]

                \end{large}

                \textbf{Categorical Cross-Entropy Function:}

                \begin{large}

                    \[ CCEL = - \frac{1}{N} \sum_{i=1}^N \sum_{j=1}^C (y_{i,j} \cdot log(p_{i,j})) \]

                \end{large}

                \noindent Where N := number of rows \newline
                C := number of classes \newline
                y := indicator if class label j is the correct clasification for observation i \newline
                p := predicted probability observation i is of class j
